A Convolutional Neural Network for Outsole Recognition\tnoteref{t1}

### 2019-06-03 paper/paper.Rnw
Update intro and talk about feature detectors

In this paper, we discuss an alternate approach that uses an additional model layer to transfer the feature vector output into a vector of probabilities representing the detection of geometric elements in shoe tread images. Working within the feature space used by forensic examiners allows us to augment human-identified features with model output, assessing similarity of different shoe images on a feature set that is explicitly relevant to the domain."
\svp{Add 1-2 sentences connecting feature sets to the goal of reducing workload - possibly cite solemate as an example of manual feature identification}

\section{Materials and Methods}
\svp{Short bit (1-2 sentences) about low-level feature detectors and why they fail with this problem.}

\svp{Motivate the use of CNNs directly - complex model, higher level of feature aggregation. Mention pre-trained networks to transition.}

>>>>

In this paper, we discuss an alternate approach that uses an additional model layer to transfer the feature vector output into a vector of probabilities representing the detection of geometric elements in shoe tread images. Working within the feature space used by forensic examiners allows us to augment human-identified features with model output, assessing similarity of different shoe images on a feature set that is explicitly relevant to the domain.
%\svp{Add 1-2 sentences connecting feature sets to the goal of reducing workload - possibly cite solemate as an example of manual feature identification}
\mt{In addition, much of the existing software to assess geometric similarity, such as SoleMate \fix{how to cite software?}, requires examiners to identify and label features manually; automating classification within the currently used feature space could reduce or eliminate the need for such manual classification.}

\section{Materials and Methods}
%\svp{Short bit (1-2 sentences) about low-level feature detectors and why they fail with this problem.}
\mt{Shoeprint evidence from crime scenes is most commonly collected in the form of a photograph, so any useful method to automatically identify outsole characteristics must take the form of an image analysis task. There are a number of methods that may be employed to identify shapes and features in an image, such as generalized Hough transformations \citep{ballardGeneralizingHoughTransform1981} and other low-level feature extraction methods aimed at detecting specific shapes, such as edges, corners, blobs, or ridges \citep[Ch 15]{machineVision}. While these methods are useful in identifying these specific features at a low level, they are very computationally intensive and only identify features on a very small scale; as a result, they cannot reliably identify large geometric shapes like those that may be found in an outsole image.}
\fix{Include? (my gut says its not necessary):} \mt{These classical methods are also extremely sensitive to lighting or color changes. As one or more models would be required for each geometric shape, classical methods would require use of additional machine learning models such as random forests to aggregate low-level features into functional geometries found on outsoles.}

%\svp{Motivate the use of CNNs directly - complex model, higher level of feature aggregation. Mention pre-trained networks to transition.}
Convolutional neural networks (CNNs) are widely recognized as superior for novel image classification. CNNs are a form of artificial neural network which make use of the image convolution operator used by many low-level feature extraction methods.%\footnote{The Google Trends interest graph for CNNs and computer vision shows a massive increase in convolutional neural networks between mid-2014 and 2018 \url{https://trends.google.com/trends/explore?date=2010-01-01\%202019-02-18&q=convolutional\%20neural\%20network,computer\%20vision}}.
As CNNs have evolved, their architecture has become more complex, but the fundamental reliance on the image convolution operator sets CNNs apart from other artificial neural networks \citep{guRecentAdvancesConvolutional2018}. CNNs have deep architectures that can be trained to identify complex patterns, but they are structurally similar to the human visual architecture and output binary or probabilistic predictions for given labels that are readily interpretable. As CNNs make use of labeled training data, the predictions generated are for features which are similar to those identified by humans, resulting in models with greater face validity. Once a CNN is trained, it is relatively fast and easy to apply the model to new images and obtain classifications.
